Forecasting Presidential Candidate Support during the Final Week before Election Day

### 2024-11-04 paper/paper.qmd
Added the abstract, fixed the subtitle, added a reference of the modelsummary package, added cross-references.

---
title: "Forecasting Presidential Candidate Support during the Final Week before Election Day"
subtitle: "In the final week leading up to the 2024 US election, Kamala Harris is expected to lead the polls with 48.03% of support with Donald Trump following close behind with 47.98% of support."
author: 
  - Tianning He
  - Julia Lee
  - Shuangyuan Yang
thanks: "Code and data are available at: https://github.com/JuliaJLee/Forecasting_US_Election_2024.git"
date: today
date-format: long
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Introduction

@@

To simulate, test, download, clean, and model data throughout this analysis, the statistical programming language R was used [@citeR]. Specific libraries that assisted the analysis include `tidyverse` [@tidyverse], `dplyr` [@dplyr],`tinytex` [@tinytex], `ggplot2` [@ggplot2], `knitr` [@knitr], `arrow` [@arrow], `here` [@here], and `testthat` [@testthat].

@@

*Code that runs through the following steps can be found in the repository linked on page 1.*

@@

@@

## Copy of the survey questions

>>>>

---
title: "Forecasting Presidential Candidate Support during the Final Week before Election Day"
subtitle: "In the final week before the election, Harris is expected to lead the polls with 48.03% of support with Trump following close behind with 47.98% of support."
author: 
  - Tianning He
  - Julia Lee
  - Shuangyuan Yang
thanks: "Code and data are available at: https://github.com/JuliaJLee/Forecasting_US_Election_2024.git"
date: today
date-format: long
abstract: "With the Presidential General Election Polls data provided by 538, the present analysis builds a model to forecast the percentage of support Harris and Trump will receive during the week before election day. By pooling poll data by week through the use of seasonal indexes and a linear model, this analysis forecasts that prior to election day, Harris will receive 48.03% of support and Trump will receive 47.98% of support from the polls. This suggests that the election will be a very tight race that may come down to the outcomes of the swing states. Through this, this analysis further examines the impacts of pooling poll data and discusses how improvements may be made to its model."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Introduction {#sec-page1}

The recent US presidential election has been a hot topic of discussion, and as the election approaches, speculation about the final outcome has begun. Accurately predicting the outcome has become a focus of attention for political analysts, the media, and the public. Polling organizations such as YouGov have conducted surveys that provide a deeper look into voters' intentions and preferences, which can help to forecast potential election results. However, people’s intentions, preferences, and opinions can be influenced by different social contexts [@intro], meaning that voters’ opinions on who should be the next president of their country can change over time.

With this, the objective of the present analysis is to build a model that forecasts the percentage of support Kamala Harris and Donald Trump will receive in the final week before the election, while accounting for variations across time. The opinions and preferences of voters’ are what this analysis is interested in. By using seasonal indexes along with a regression model, this analysis seeks to measure and approximate voters’ opinions to generate more precise forecasts that reflect their preferences.

This analysis first aims to understand how the percentage of support for both candidates has changed over time along with how the percentage of support for one candidate has changed relative to the other. Then, as a means to observe how the model affects the data, this analysis examines how the percentages of support predicted by the linear model compare to the weighted averages over time. Finally, the model is used to predict the percentage of support Harris and Trump will receive in the final week before the election. Additionally, this analysis considers the accuracy of the linear model's predictions and investigates the impact of the “pooling polls” method on election predictions. This provides a deeper awareness of how the complexity that surrounds having an opinion can be translated into data that ultimately predicts real-world election results.

As a result, this analysis finds that Harris will lead the polls with 48.03% of support with Trump close behind with 47.98% of support, suggesting that this election will be a tight race that may need to be determined by the outcomes of the swing states. These findings further highlight important considerations for how to improve the ways in which poll data is interpreted and different prediction methods for future elections. The following expands on the ideas above by presenting a description of the poll data for this analysis (@sec-data), a more in-depth explanation of model and how it is used (@sec-model), an overview of the results (@sec-results), and a discussion of the overall analysis (@sec-discussion).

@@

To simulate, test, download, clean, and model data throughout this analysis, the statistical programming language R was used [@citeR]. Specific libraries that assisted the analysis include `tidyverse` [@tidyverse], `dplyr` [@dplyr],`tinytex` [@tinytex], `ggplot2` [@ggplot2], `knitr` [@knitr], `arrow` [@arrow], `here` [@here], `testthat` [@testthat], and `modelsummary` [@modelsummary].

@@

Code that runs through the following steps can be found in the repository linked on page 1 (@sec-page1).

@@

A copy of the survey questions can be found in @sec-copy.

@@

## Copy of the survey questions {#sec-copy}
